Double White Dwarf Mergers with CASTRO\\ I. Methodology and Code Verification

### 2015-01-01 papers/wdmerger_I/paper.tex
Add DD acronym

Type Ia supernovae (SNe Ia) are currently some of the most exciting events to study in astrophysics. These bright, brief pulses of light in the distant universe have led to a number of important discoveries in recent years, including the discovery of the accelerated expansion of the universe \citep{perlmutter1999,riess1998}. However, their origin is shrouded in mystery. It has long been expected that these events arise from the thermonuclear explosions of white dwarfs \citep{hoyle_fowler:1960}, but the cause of these explosions is uncertain. In particular, it is not clear what process causes the temperatures in these white dwarfs to become hot enough for explosive burning of its constituent nuclei. The model favored initially by the community was the so-called single-degenerate (SD) model \citep{whelan_iben:1973}. Accretion of material from a companion star such as a red giant would cause the star to approach the Chandrasekhar mass, and in doing so the temperature and density in the center would become sufficient for thermonuclear fusion to proceed. However, in recent years many alternative progenitor models have been discussed. A leading candidate for explaining some or most of these explosions is the double-degenerate model, in which two white dwarfs merge and the merged object reaches the conditions necessary for a thermonuclear ignition \citep{ibentutukov:1984,webbink:1984}. Another is the double detonation scenario, where accretion of material onto a sub-Chandrasekhar white dwarf would lead to a detonation inside the accreted envelope, and this would send a compressional wave into the core of the star that would trigger a secondary detonation. A recent review of the progenitor models can be found in \citet{hillebrandt:2013}.

>>>>

Type Ia supernovae (SNe Ia) are currently some of the most exciting events to study in astrophysics. These bright, brief pulses of light in the distant universe have led to a number of important discoveries in recent years, including the discovery of the accelerated expansion of the universe \citep{perlmutter1999,riess1998}. However, their origin is shrouded in mystery. It has long been expected that these events arise from the thermonuclear explosions of white dwarfs \citep{hoyle_fowler:1960}, but the cause of these explosions is uncertain. In particular, it is not clear what process causes the temperatures in these white dwarfs to become hot enough for explosive burning of its constituent nuclei. The model favored initially by the community was the so-called single-degenerate (SD) model \citep{whelan_iben:1973}. Accretion of material from a companion star such as a red giant would cause the star to approach the Chandrasekhar mass, and in doing so the temperature and density in the center would become sufficient for thermonuclear fusion to proceed. However, in recent years many alternative progenitor models have been discussed. A leading candidate for explaining some or most of these explosions is the double-degenerate (DD) model, in which two white dwarfs merge and the merged object reaches the conditions necessary for a thermonuclear ignition \citep{ibentutukov:1984,webbink:1984}. Another is the double detonation scenario, where accretion of material onto a sub-Chandrasekhar white dwarf would lead to a detonation inside the accreted envelope, and this would send a compressional wave into the core of the star that would trigger a secondary detonation. A recent review of the progenitor models can be found in \citet{hillebrandt:2013}.
